\Large \bf Uncovering Distributed System Bugs during Testing (not Production!)

### 2015-09-19 evaluation.tex
lots of updates to address Ally's comments

Environmental modeling is a core activity of using \psharp. It is required for \emph{closing the environment} of a system-under-test and making it amenable to systematic testing. Table~\ref{tab:stats} presents program statistics for our three case studies. We report: lines of code for the system-under-test (\#LoC); number of bugs found in the system (\#B); lines of \psharp code for the environmental model (\#LoC); number of machines (\#M); number of state transitions (\#ST); and number of action handlers (\#AH).

Modeling the environment of the Extent Manager in the Azure Storage vNext system required approximately 2 weeks of part-time developing. The \psharp model for testing this system is the smallest (in lines of code) from all three case studies. This was because the modeling effort was targeting the particular liveness bug that was haunting the developers of vNext. We are currently in the process of modeling other components of vNext, such as a ChainReplication and a Paxos system.

Developing the Live Migration Table and its \psharp test harness took about 5 weeks. The test harness was developed in parallel with the actual system.
This is in contrast with the other two case studies discussed in this paper, where the modeling 
activity occurred independently and at a later stage of the development process.

Modeling Fabric required approximately $5$ person months. Although this is a
significant amount of time, it is a one time effort, which only needs incremental refinement with each release.

@@

Using \psharp we managed to uncover more than 10 serious bugs in our case studies. As discussed earlier in the paper, these bugs were hard to find with traditional testing techniques, but \psharp managed to uncover them and reproduce them in a small setting. According to the developers, the traces of \psharp were useful, as it allowed them to understand the source of the bug and fix it in a timely manner. After the developers fixed all the discovered bugs, we optionally reintroduced them one-by-one so that we can evaluate the effectiveness of different \psharp systematic testing strategies in finding these bugs.

@@

For the MigratingTable case study, the first 7 bugs in Table~\ref{tab:testing} were discovered using a \psharp test harness based on nondeterministically generated, but controlled by \psharp, input history. For each of these found bugs, we manually reviewed one of the bug traces to confirm if it reflected the expected bug. The remaining 4 bugs (denoted with $\diamond$) were not caught with our default test harness in the 100,000 iterations. We believe this is due to unlucky random choices of inputs and schedules by the \psharp testing engine. We wrote a custom \psharp test harness for each of these bugs, which allowed \psharp to quickly reproduce them. The design of these custom harnesses was influenced by the knowledge of these bugs; our objective here was to simply see if \psharp can reproduce these known bugs.

>>>>

Environmental modeling is a core activity of using \psharp. It is required for \emph{closing the environment} of a system to make it amenable to systematic testing. Table~\ref{tab:stats} presents program statistics for the three case studies. The columns under ``System'' refer to the real system-under-test, while the columns under ``\psharp model'' refer to the \psharp test harness. We report: lines of code for the system-under-test (\#LoC); number of bugs found in the system-under-test (\#B); lines of \psharp code for the test harness (\#LoC); number of machines (\#M); number of state transitions (\#ST); and number of action handlers (\#AH).

Modeling the environment of the Extent Manager in the Azure Storage vNext system required approximately two weeks of part-time developing. The \psharp model for testing this system is the smallest (in lines of code) from the three case studies. This was because this modeling exercise aimed to reproduce the particular liveness bug that was haunting the developers of vNext.

Developing the Live Migration Table and its \psharp test harness took about five weeks. The test harness was developed in parallel with the actual system. This is in contrast with the other two case studies discussed in this paper, where the modeling activity occurred independently and at a later stage of the development process.

Modeling Fabric required approximately five person months. Although this is a significant amount of time, it is a one-time effort, which only needs incremental refinement with each release.

@@

Using \psharp we managed to uncover 12 serious bugs in our case studies \PDComment{update for Fabric}. As discussed earlier in the paper, these bugs were hard to find with traditional testing techniques, but \psharp managed to uncover them and reproduce them in a small setting. According to the developers, the traces of \psharp were useful, as it allowed them to understand the source of the bug and fix it in a timely manner. After the developers fixed all the discovered bugs, we optionally reintroduced them one-by-one to evaluate the effectiveness of different \psharp systematic testing strategies in finding these bugs.

@@

For the MigratingTable case study, the first seven bugs in Table~\ref{tab:testing} were discovered using a \psharp test harness based on nondeterministically generated, but controlled by \psharp, input history. For each of these found bugs, we manually reviewed one of the bug traces to confirm if it reflected the expected bug. The remaining four bugs (denoted with $\diamond$) were not caught with our default test harness in the 100,000 iterations. We believe this is due to unlucky random choices of inputs and schedules by the \psharp testing engine. We wrote a custom \psharp test harness for each of these bugs, which allowed \psharp to quickly reproduce them. The design of these custom harnesses was influenced by the knowledge of these bugs; our objective here was to simply see if \psharp can reproduce these known bugs.

### 2015-09-19 experience.tex
lots of updates to address Ally's comments

The modeling and testing approach described in earlier sections of the paper is not specific to the vNext system. \psharp is a generic framework. We showcase this capability by presenting developer experience of using \psharp to model and test two other distributed storage systems used in production in Microsoft.

@@

The main challenge behind testing MigratingTable is that there are many possible input histories and that the system is highly concurrent. The developers could have tested specific input histories, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult to foresee interactions between different parts of the code.

\subsubsection{Testing the MigratingTable library}

Towards testing the MigratingTable library, the developers wrote an in-memory reference implementation of the \texttt{IChainTable} interface, called \texttt{SpecTable}, which can be used for comparing the output of MigratingTable on an arbitrary input history. This enabled sampling from a distribution that was defined over all possible input histories within certain bounds. The \psharp systematic testing engine automatically takes care of enumerating different input histories.

@@

The MigratingTable was instrumented to report the intended \emph{linearization point} of each input call. Specifically, after each input call completes, MigratingTable reports whether that call was the linearization point, which may depend on the result of the call.  This makes it possible to check the correctness property as the model executes.

The \psharp environmental model of MigratingTable consists of a \texttt{Tables} machine containing the old, new and reference table implementations; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}).

Each \texttt{Service} machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the \texttt{Tables} machine. When MigratingTable reports the linearization point of an input call, the \texttt{Service} machine sends that input call to the reference table.  When an input call completes, the \texttt{Service} machine checks that the results from the MigratingTable and the reference table agree.

@@

Azure Service Fabric\footnote{\url{azure.microsoft.com/campaigns/service-fabric/}} (or Fabric for short) is a platform and API for creating reliable services that execute on a cluster of machines. 
%The developer writes a service that receives requests (e.g.\ from some client program via HTTP requests) and mutates its state based on these requests. 
In order to make the user-written service \emph{reliable}, Fabric launches several \emph{replicas} (copies) of the service, where each replica runs as a separate process on a different node in the cluster.
%\PTComment{Cut: description of primary and secondaries, and electing a new primary.}
%One replica is selected to be the \emph{primary} which serves client requests; the rest are \emph{secondaries}. The primary replicates state changes to the secondaries by sending \emph{replication requests} so that all replicas eventually have the same state. If the primary fails (e.g.\ if the node on which the primary is running crashes), Fabric elects one of the secondaries to be the new primary and launches another secondary; the new secondary will receive a full or partial copy (depending on whether persistent storage is used) of the state of the new primary in order to ``catch up'' with the other secondaries. Fabric provides a name-resolution service so that clients can always find the current primary.
The state of the service is replicated from the \emph{primary} replica to the other \emph{secondary} replicas for redundancy.
User-written Fabric services are complex asynchronous and distributed applications and are, thus,
challenging to test.
% They are interesting targets for systematic testing with \psharp{}.

@@

\psharp's systematic testing was very helpful in debugging the model itself.
Without systematic testing, we feel even catching bugs in the model 
would have been difficult.

@@

The main system that we tested is \emph{CScale}~\cite{faleiro2012},
a big data-stream processing system. In prior work~\cite{deligiannis2015psharp}, 
we used an earlier version of the Fabric model and reported bugs in sample serives.
Supporting CScale required significant additions to the model, making it much more
feature complete.
CScale chains multiple Fabric services, which communicate via remote procedure calls (RPCs).
To close the system, we modeled RPCs 
by sending and receiving \psharp{} events.
This was enough to converted a distributed system
that uses both Fabric and its own
network communication protocol
% that runs on the Fabric platform
% and 
% uses network communication
into 
a closed single process system. 
% that uses multiple threads,
% contains
% the Fabric model
% and the CScale services,
% and does not use network communication.
%\PTComment{Cut: we had to remove static fields.}
% Additional changes were needed to
% remove certain static (per-process) fields,
% as these were inadvertently
% being shared between services
% after making the system a single process.
A key challenge in our work
was to test CScale despite the fact that it
uses various synchronous and asynchronous APIs
other than \psharp{}.
This work is still in-progress.
However, we were able to find a \texttt{NullReferenceException}
bug in CScale by running it on our model. The bug has been
communicated to the developers of CScale, but we are still awaiting a
confirmation.

>>>>

The modeling and testing approach described in earlier sections of the paper is not specific to the vNext system. \psharp is a generic framework for .NET distributed systems. We showcase this capability by presenting developer experience of using \psharp to model and test two other systems used in production inside Microsoft.

@@

The main challenge behind testing MigratingTable is that there are many possible input histories and that the system is highly concurrent. The developers could have tested specific input histories, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code.

\textbf{Modeling and testing}.
Towards testing the MigratingTable library, the developers wrote an in-memory reference implementation of the \texttt{IChainTable} interface, called \texttt{SpecTable}, which can be used for comparing the output of MigratingTable on an arbitrary input history. This enabled sampling from a distribution that was defined over all possible input histories within certain bounds. The \psharp systematic testing engine automatically takes care of enumerating different input histories. The MigratingTable was instrumented to report the intended \emph{linearization point} of each input call. Specifically, after each input call completes, MigratingTable reports whether that call was the linearization point, which may depend on the result of the call.  This makes it possible to check the correctness property as the model executes.

@@

The \psharp test harness of MigratingTable consists of a \texttt{Tables} machine containing the old, new and reference table implementations; a collection of \texttt{Service} machines containing identically configured MigratingTables; and a \texttt{Migrator} machine that performs the background migration (see Figure~\ref{fig:mockedmigration}). Each \texttt{Service} machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the \texttt{Tables} machine. When MigratingTable reports the linearization point of an input call, the \texttt{Service} machine sends that input call to the reference table.  When an input call completes, the \texttt{Service} machine checks that the results from the MigratingTable and the reference table agree.

@@

Azure Service Fabric\footnote{\url{azure.microsoft.com/campaigns/service-fabric/}} (or Fabric for short) is a platform and API for creating reliable services that execute on a cluster of machines. 
%The developer writes a service that receives requests (e.g.\ from some client program via HTTP requests) and mutates its state based on these requests. 
In order to make the user-written service \emph{reliable}, Fabric launches several \emph{replicas} (copies) of the service, where each replica runs as a separate process on a different node in the cluster.
%\PTComment{Cut: description of primary and secondaries, and electing a new primary.}
%One replica is selected to be the \emph{primary} which serves client requests; the rest are \emph{secondaries}. The primary replicates state changes to the secondaries by sending \emph{replication requests} so that all replicas eventually have the same state. If the primary fails (e.g.\ if the node on which the primary is running crashes), Fabric elects one of the secondaries to be the new primary and launches another secondary; the new secondary will receive a full or partial copy (depending on whether persistent storage is used) of the state of the new primary in order to ``catch up'' with the other secondaries. Fabric provides a name-resolution service so that clients can always find the current primary.
The state of the service is replicated from the \emph{primary} replica to the other \emph{secondary} replicas for redundancy.
User-written Fabric services are complex asynchronous and distributed applications, and are thus challenging to test.
% They are interesting targets for systematic testing with \psharp{}.

@@

\psharp's systematic testing was very helpful in debugging the model itself.
Without systematic testing, we feel even catching bugs in the model 
would have been difficult.
\PDComment{See Ally's feedback and maybe try to address this as we now have some extra space in the paper: ``\psharp systematic testing was very helpful in debugging the model itself.'' this short paragraph is very intriguing, and begs a little more elaboration about what sorts of bugs you typically found in the model.}

@@

The main system that we tested is \emph{CScale}~\cite{faleiro2012},
a big data-stream processing system. In prior work~\cite{deligiannis2015psharp}, 
we used an earlier version of the Fabric model and reported bugs in sample serives.
Supporting CScale required significant additions to the model, making it much more
feature-complete.
CScale chains multiple Fabric services, which communicate via remote procedure calls (RPCs).
To close the system, we modeled RPCs 
by sending and receiving \psharp{} events.
This was enough to converted a distributed system
that uses both Fabric and its own
network communication protocol
% that runs on the Fabric platform
% and 
% uses network communication
into 
a closed single process system. 
% that uses multiple threads,
% contains
% the Fabric model
% and the CScale services,
% and does not use network communication.
%\PTComment{Cut: we had to remove static fields.}
% Additional changes were needed to
% remove certain static (per-process) fields,
% as these were inadvertently
% being shared between services
% after making the system a single process.
A key challenge in our work
was to test CScale despite the fact that it
uses various synchronous and asynchronous APIs
other than \psharp{}.
This work is still in-progress.
However, we were able to find a \texttt{NullReferenceException}
bug in CScale by running it on our model. The bug has been
communicated to the developers of CScale, but we are still awaiting a
confirmation.

### 2015-09-19 methodology.tex
update to remove inbalanced subsubsections
lots of updates to address Ally's comments

The system-under-test, the real ExtMgr in vNext, is wrapped inside the \texttt{ExtentManager} \psharp machine in the testing harness.

\subsubsection{Internals of the real Extent Manager}

@@

Inside the real ExtMgr, there are two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} maintains the mapping records from extents to their hosting ENs. It is updated upon the periodic sync reports from the ENs. Recall that the sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update extent manager's possible out-of-date view of the EN with the ground truth. The EN map records the latest heartbeat time from every EN.

@@

\subsubsection{Intercepting messages between machines}
\label{sec:method:model:dispath}

@@

The real ExtMgr uses a network engine to send messages to the ENs. The testing harness mocks the original network engine in vNext and overrides its interface. In this way, the mocked network engine intercepts all outbound messages and relays them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines. As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}. Conceivably, the mocked network engine could leverage the non-determinism support in \psharp and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable. \CHComment{only if non-determinism is already covered in the previous section.}

@@

\subsubsection{Specifics of the ExtentManager machine}

Figure~\ref{fig:wrap_target} shows code snippets from the \texttt{ExtentManager} \psharp machine. \texttt{ExtentManager} is a thin wrapper of the real ExtMgr. The mocked network engine replaces the real one in ExtMgr, intercepts all the outbound messages from ExtMgr and relays them to \texttt{TestingDriver}.

@@

Messages coming from the ENs do {\em not} go through the mocked network engine. They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the internal ExtMgr with \texttt{extMgr.ProcessSingleMessage}. The benefit of this approach is that ExtMgr is unaware of the testing harness and behaves as if it running in a real distributed environment and communicating with real ENs.

All internal timers of ExtMgr are disabled, because it is crucial to delegate all non-determinism to \psharp. Instead, the EN expiration loop and the extent repair loop are driven by external timers modeled in \psharp. Note that system correctness should {\em not} hinge on the frequency of any individual timer. Hence, the \psharp systematic testing engine has the complete freedom to schedule arbitrary interleavings between the timers.

@@

The \texttt{ExtentNode} machine is a much simplified version of the original EN. It omits the details of a real EN, mocks only logic necessary for the testing scenarios, and keeps the mocked logic as simple as possible. The logic here includes: repairing an extent from its replica, generating sync reports periodically, and sending heartbeat messages periodically.

The testing harness leverages components in vNext whenever appropriate. Here, the \texttt{ExtentNode} machine re-uses the \texttt{ExtentCenter} data structure, an internal component of a real EN used for extent bookkeeping.

@@

The \texttt{TestingDriver} machine drives two testing scenarios. In the first scenario, it launches one \texttt{ExtentManager} and three \texttt{ExtentNode} machines, with a single extent on one of the ENs. It waits for the extent to be replicated at the other ENs. In the second scenario, it fails one of the \texttt{ExtentNode} machines and launches a new one. It waits for the extent to be repaired on the new EN. Code snippets are skipped due to space constraint.

@@

\texttt{RepairMonitor} is a monitor machine (see Section~\ref{sec:bg:bugs}) that transitions between a cold state and a hot state. Whenever an extent node fails, \texttt{RepairMonitor} is notified. As soon as the number of extent replicas falls below a specified target, \texttt{RepairMonitor} transitions into the hot \emph{repairing} state, where the missing replica is being repaired. Whenever a replica is repaired, the \texttt{RepairMonitor} machine is also notified. It transitions into the cold \emph{repaired} state, when the replica number reaches the target again, as illustrated in Figure~\ref{fig:monitor}.

%It checks the property that the number of extent replica {\em eventually} reaches the target.

In the extent repair testing scenarios, \texttt{RepairMonitor} checks that it should {\em always eventually} end up in the cold repaired state. Otherwise, the \texttt{RepairMonitor} machine is stuck in the hot repairing state for {\em infinitely} long. This indicates that the corresponding execution sequence results in an extent replica never being repaired, which is a liveness bug!

@@

It took only tens of seconds before the testing harness reported the first occurrence of a liveness bug. Upon examining the debug trace, the vNext developers were able to confirm the bug.

However, the trace didn't include enough details, which prevented the developers from identifying a root cause. Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debug outputs in each iteration. After several iterations, the developers were able to pinpoint the exact culprit and immediately proposed a solution to fix the bug. Once the proposed solution was implemented, the developers again ran the testing harness, which reported no more bug for $100,000$ iterations in tens of minutes.

@@

Here is one particular execution sequence resulting in the bug. $i)$ EN$_0$ fails and is detected by the EN expiration loop; $ii)$ EN$_0$ is removed from the EN map; $iii)$ the extent center is updated and the replica count drops from 3, the target, to 2; $iv)$ \texttt{ExtMgr} receives a sync report from EN$_0$; $v)$ the extent center is updated and the replica count increases from 2 to 3 again. This is problematic because, on one hand, the replica count is equal to the target, so the extent repair loop never schedules any repair task. On the other hand, there are only two true replicas in the system, one fewer than the target. This execution sequence leads to one replica missing. Conceivably, repeating this another two times would result in all replicas missing, while \texttt{ExtMgr} still thinks all replicas are healthy. If deployed in production, such bug would have caused a very serious incident of customer data loss.

The culprit is in $iv)$, where \texttt{ExtMgr} receives a sync report from EN$_0$ after deleting the EN. This may occur in \psharp because of arbitrary message interleaving. It may also occur, albeit much less frequently, in the stress testing due to messages being delayed in the network. This explains why the bug only occurs from time to time in the stress testing and also takes long execution to trigger. In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit identified promptly, and the fix solution tested effectively and thoroughly, all of which should vastly increase the productivity of distributed storage system development.

>>>>

The real ExtMgr in vNext, which is our system-under-test, is wrapped inside the \texttt{ExtentManager} \psharp machine in the testing harness.

@@

\textbf{Internals of the real Extent Manager}.
Inside the real ExtMgr (see Figure~\ref{fig:extentmanager}), there are two data structures related to extent replication and repair: \texttt{ExtentCenter} and \texttt{ExtentNodeMap}. The \texttt{ExtentCenter} maintains the mapping records from extents to their hosting ENs. It is updated upon the periodic sync reports from the ENs. Recall that the sync report from a particular EN lists all the extents stored at the EN. Its purpose is to update extent manager's possible out-of-date view of the EN with the ground truth. The EN map records the latest heartbeat time from every EN.

@@

\textbf{Intercepting messages between machines}.
The real ExtMgr uses a network engine to send messages to the ENs. The testing harness mocks the original network engine in vNext and overrides its interface. In this way, the mocked network engine intercepts all outbound messages and relays them to the \texttt{TestingDriver} machine, which is responsible for dispatching the messages to the corresponding \texttt{ExtentNode} machines. As shown in Figure~\ref{fig:enginecode}, the mocked network engine intercepts the outbound messages from the ExtMgr and invokes \texttt{PSharpRuntime.Send(...)} to asynchronously relay the messages to \texttt{TestingDriver}. Conceivably, the mocked network engine could leverage the nondeterminism support in \psharp and choose to drop the messages in a non-deterministic fashion, in case emulating message loss is desirable.

@@

@@

\textbf{Specifics of the ExtentManager machine}.
Figure~\ref{fig:wrap_target} shows code snippets from the \texttt{ExtentManager} \psharp machine. \texttt{ExtentManager} is a thin wrapper of the real ExtMgr. The mocked network engine replaces the real one in ExtMgr, intercepts all the outbound messages from ExtMgr and relays them to \texttt{TestingDriver}.

@@

Messages coming from the ENs do \emph{not} go through the mocked network engine. They are delivered to the \texttt{ExtentManager} machine directly and trigger an action that invokes the messages on the internal ExtMgr with \texttt{extMgr.ProcessSingleMessage}. The benefit of this approach is that ExtMgr can be tested without modifying its code; ExtMgr is simply unaware of the testing harness and behaves as if it running in a real distributed environment and communicating with real ENs.

System correctness should \emph{not} hinge on the frequency of any individual timer. Instead, all nondeterminism due to timing-related events must be delegated to \psharp. To achieve this, all timers inside ExtMgr are disabled, and the EN expiration loop and the extent repair loop are driven instead by timers modeled in \psharp, an approach that was used in previous work~\cite{desai2015building}. The \psharp timers send nondeterministic timeout events that are controlled by the \psharp runtime. Hence, the \psharp testing engine has the freedom to schedule arbitrary interleavings between timeout events and all other regular system events.

@@

The \texttt{ExtentNode} machine is a simplified version of the original EN. The machine omits much of the complex details of a real EN, and only mocks the logic necessary for the testing scenarios. This mocked logic includes: repairing an extent from its replica, and sending sync reports and heartbeat messages periodically to the \texttt{ExtentManager} machine.

The testing harness leverages components of Azure Storage vNext whenever it is appropriate. For example, the \texttt{ExtentNode} machine re-uses the \texttt{ExtentCenter} data structure, a component that is used inside a real EN for extent bookkeeping.

@@

The \texttt{TestingDriver} machine drives two testing scenarios. In the first scenario, it launches one \texttt{ExtentManager} and three \texttt{ExtentNode} machines, with a single extent on one of the ENs. It waits for the extent to be replicated at the other ENs. In the second scenario, it fails one of the \texttt{ExtentNode} machines and launches a new one. It waits for the extent to be repaired on the new EN. We omit illustrative code snippets due to space constraints.

@@

\texttt{RepairMonitor} is a liveness monitor machine (see Section~\ref{sec:bg:bugs}) that transitions between a cold and a hot state. Whenever an extent node fails, \texttt{RepairMonitor} is notified. As soon as the number of extent replicas falls below a specified target (three replicas in the current test harness), \texttt{RepairMonitor} transitions into the hot \emph{repairing} state, where the missing replica is being repaired. Whenever a replica is repaired, the \texttt{RepairMonitor} machine is also notified. It transitions into the cold \emph{repaired} state, when the replica number reaches again the target, as illustrated in Figure~\ref{fig:monitor}.

In the extent repair testing scenarios, \texttt{RepairMonitor} checks that it should {\em always eventually} end up in the cold repaired state. Otherwise, the \texttt{RepairMonitor} machine is stuck in the hot repairing state for {\em infinitely} long. This indicates that the corresponding execution sequence results in an extent replica never being repaired, which is a liveness bug.

@@

It took less than ten seconds for the \psharp testing engine to report the first occurrence of a liveness bug in vNext (see Section~\ref{sec:eval}). Upon examining the debug trace, the vNext developers were able to confirm the bug.

However, the original \psharp trace did not include sufficient detail to allow the developers to identify the root cause of the problem. Fortunately, running the test harness took very little time, so the developers were able to quickly iterate and add more refined debug outputs in each iteration. After several iterations, the developers were able to pinpoint the exact culprit and immediately propose a solution for fixing the bug. Once the proposed solution was implemented, the developers ran again the test harness. No bugs were found during 100,000 iterations, a process that required only tens of minutes.

@@

The following is one particular execution sequence resulting in this liveness bug: (i) EN$_0$ fails and is detected by the EN expiration loop; (ii) EN$_0$ is removed from the EN map; (iii) the extent center is updated and the replica count drops from 3 (which is the target) to 2; (iv) ExtMgr receives a sync report from EN$_0$; (v) the extent center is updated and the replica count increases again from 2 to 3. This is problematic since the replica count is equal to the target, which means that the extent repair loop will never schedule any repair task. At the same time, there are only two \emph{true} replicas in the system, which is one less than the target. This execution sequence leads to one replica missing; repeating this process two more times would result in all replicas missing, but ExtMgr still thinking that all replicas are healthy. In production, such a bug can cause a very serious incident of customer data loss.

The culprit is in (iv), where ExtMgr receives a sync report from EN$_0$ after deleting the EN. This may occur in \psharp due to arbitrary interleaving of events. It may also occur, albeit much less frequently, during stress testing due to messages being delayed in the network. This explains why the bug only occurs from time to time during stress testing and requires long executions to manifest. In contrast, \psharp allows the bug to manifest quickly, the developers to iterate rapidly, the culprit to be identified promptly, and the fix to be tested effectively and thoroughly, all of which have the potential to vastly increase the productivity of distributed storage system development.

### 2015-09-19 motivation.tex
update to remove inbalanced subsubsections
lots of updates to address Ally's comments

\item \emph{Integration testing}, in which an ExtMgr in launched together with multiple ENs. An EN failure is subsequently injected. These tests check that the affected extents are eventually repaired.

\item \emph{Stress testing}, in which an ExtMgr in launched together with multiple ENs and multiple extents. The test keeps repeating the following process: injecting an EN failure, launching a new EN and checking that the affected extents are eventually repaired.
\end{enumerate}

Despite of the extensive testing efforts, the vNext developers have been plagued by what appears to be an elusive bug in the ExtMgr logic. All the unit test and integration test suites successfully pass every single time. However, the stress test suite fails {\em from time to time} after very long executions, manifested as the replicas of some extents remain missing while never being repaired. The bug has proven to be difficult to identify, reproduce and troubleshoot. First, it takes very long executions to trigger. Second, an extent not being repaired is {\em not} a property that can be easily verified. In practice, the developers rely on very large time-out period to detect the bug. Finally, by the time that the bug is detected, very long execution traces have been collected, which makes manual inspection tedious and ineffective.

>>>>

\item \emph{Integration testing}, in which an ExtMgr is launched together with multiple ENs. An EN failure is subsequently injected. These tests check that the affected extents are eventually repaired.

\item \emph{Stress testing}, in which an ExtMgr is launched together with multiple ENs and multiple extents. The test keeps repeating the following process: injecting an EN failure, launching a new EN and checking that the affected extents are eventually repaired.
\end{enumerate}

\noindent
Despite of the extensive testing efforts, the vNext developers have been plagued by what appears to be an elusive bug in the ExtMgr logic. All the unit test and integration test suites successfully pass every single time. However, the stress test suite fails {\em from time to time} after very long executions, manifested as the replicas of some extents remain missing while never being repaired. The bug has proven to be difficult to identify, reproduce and troubleshoot. First, it takes very long executions to trigger. Second, an extent not being repaired is {\em not} a property that can be easily verified. In practice, the developers rely on very large time-out period to detect the bug. Finally, by the time that the bug is detected, very long execution traces have been collected, which makes manual inspection tedious and ineffective.

### 2015-09-19 overview.tex
lots of updates to address Ally's comments

\textbf{Safety}.
In addition to the usual assertions that can be used to write safety properties that are local to a machine, \psharp also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, which is a special machine that can receive, but not send, events. A safety monitor maintain local state that is modified in response to events received from ordinary (i.e. non-monitor) machines. This local state is used to maintain a history of the computation that is relevant to the property being specified. An erroneous global behavior is flagged via an assertion on the private state of the safety monitor. 
Thus, a monitor cleanly separates instrumentation state required for specification (inside the monitor) from program state (outside the monitor).

@@

\textbf{Liveness}.
Liveness property specifications are more difficult to encode since they are intended to capture program progress
over infinite executions.
Violation of a liveness property can only be demonstrated by an infinite execution.
Usually, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977,lamport1994temporal}.
We take a different approach and allow the programmer to write a \emph{liveness monitor}.
Similar to a safety monitor, a liveness monitor receives events but does not send them.
Unlike a safety monitor, a liveness monitor~\cite{desai2015building}
contains three types of states---\emph{ordinary}, \emph{hot}, and \emph{cold}.
A hot state denotes a point in the execution where progress is required but has not happened yet,
e.g., a node has failed but a new one has not come up yet.
A liveness monitor enters a hot state upon receiving notification of an event that requires the system
to make progress and leaves the hot state upon receiving another event notifying it of progress.
An infinite execution is erroneous if the liveness monitor is in a hot state infinitely often but in a cold state only finitely often.
Our liveness monitors can encode arbitrary temporal logic properties.

@@

The \psharp systematic testing engine will repeatedly execute a program from start to completion, each time exploring a different set of scheduling decisions, until it either reaches a bound (in number of iterations or time), or it hits a property violation. This testing process is fully automatic, has no false-positives (assuming an accurate environmental model), and can reproduce found bugs by replaying buggy schedules. After a bug is found, the engine generates a trace that represents the buggy schedule. In contrast to logs typically generating during production, this log provides a global order of all communication events and, thus, is easier to debug.

>>>>

\textbf{Safety specifications}.
In addition to the usual assertions that can be used to write safety properties that are local to a machine, \psharp also provides a way to specify global assertions by using a \emph{safety monitor}~\cite{desai2015building}, which is a special machine that can receive, but not send, events. A safety monitor maintains local state that is modified in response to events received from ordinary (non-monitor) machines. This local state is used to maintain a history of the computation that is relevant to the property being specified. An erroneous global behavior is flagged via an assertion on the private state of the safety monitor. 
Thus, a monitor cleanly separates instrumentation state required for specification (inside the monitor) from program state (outside the monitor).

@@

\textbf{Liveness specifications}.
Liveness property specifications are more difficult to encode since they are intended to capture program progress
over infinite executions.
Violation of a liveness property can only be demonstrated by an infinite execution.
Usually, a liveness property is specified via a temporal logic formula~\cite{Pnueli1977,lamport1994temporal}.
We take a different approach and allow the programmer to write a \emph{liveness monitor}.
Similar to a safety monitor, a liveness monitor receives events but does not send them.
Unlike a safety monitor, a liveness monitor~\cite{desai2015building}
contains three types of states---\emph{ordinary}, \emph{hot}, and \emph{cold}.
A hot state denotes a point in the execution where progress is required but has not happened yet,
e.g., a node has failed but a new one has not come up yet.
A liveness monitor enters a hot state upon receiving notification of an event that requires the system
to make progress and leaves the hot state upon receiving another event notifying it of progress.
An infinite execution is erroneous if the liveness monitor is in a hot state infinitely often but in a cold state only finitely often.
Our liveness monitors can encode arbitrary temporal logic properties.

@@

The \psharp systematic testing engine will repeatedly execute a program from start to completion, each time exploring a potentially different set of scheduling decisions, until it either reaches a bound (in number of iterations or time), or it hits a property violation. This testing process is fully automatic, has no false-positives (assuming an accurate environmental model), and can reproduce found bugs by replaying buggy schedules. After a bug is found, the engine generates a trace that represents the buggy schedule. In contrast to logs typically generating during production, this log provides a global order of all communication events and, thus, is easier to debug.
